SunPy - Python for Solar Physics

### 2014-01-21 2-1-Map.tex
add stuff to dataypes

\subsection{Map}
Map is a 2D spatial data type, primarily used for images of the Sun and 
inner heliosphere. It provides a wrapper around a data array (numpy 
ndarray) and gives easy access to standard meta data in the header of the 
image.
The \texttt{Map} datatype provides convenience methods for many functions 
such as, rotation and re-sampling as well as convenience visualisation 
functions.
The design of the map submodule is such that each instrument or 
detector can subclass the parent \texttt{GenericMap} class to implement 
special meta data handling or other data specific functions. Each subclass 
of \texttt{GenericMap} can register with the \texttt{Map} factory class and 
by implementing a method that returns \texttt{True} if the meta data 
matches meta data for that instrument or detector, the \texttt{Map} factory 
will automatically return an instance of the specific \texttt{GenericMap} 
subclass. As of version 0.4 SunPy has \texttt{Map} specialisations for the 
following instruments: \textit{IRIS} SJI frames, \textit{SDO/AIA} and 
\textit{HMI}, \textit{SOHO/EIT} and 	\textit{LASCO}, 
\textit{STEREO/EUVI} 	and \textit{COR}, \textit{YOHKOH/SXT}, 
\textit{RHESSI}, \textit{PROBA2/SWAP} and\textit{HINODE/XRT}.

\begin{listing}[h]
\begin{minted}{pycon}

@@

>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aia
SunPy AIAMap
---------
Observatory: SDO
Instrument:  AIA_3
Detector:    AIA
Measurement: 171
Obs Date:    2011-03-19T10:54:00.34
dt:          1.999601
Dimension:   [1024, 1024]
[dx, dy] =   [2.400000, 2.400000]

@@

\end{minted}
\caption{Example of printing information about a map at a Python interactive 
prompt.}
\label{code:aia_2}
\end{listing}

\begin{listing}[h]
\begin{minted}{python}

import sunpy.map

aiamap = sunpy.map.Map('aia_file.fits')
aia.peek()
\end{minted}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\caption{Demonstration of the \textit{AIA} map quick view plotting method.}
\label{code:aia_1}
\end{listing}

As can been seen in Listing \ref{code:aia_1} \& \ref{code:aia_2} the map 
datatype parses header information for many purposes including visualisation on 
correct physical axes and limb plotting as well as making this information 
easily available to the user.

@@

\begin{listing}[h]
\begin{minted}{python}
import matplotlib.pyplot as plt
import sunpy.map

@@

\begin{listing}[h]
\begin{minted}{python}
import sunpy.map

>>>>

\subsection{Map}
Map is a 2D spatial data type, primarily used for images of the Sun and 
inner heliosphere. It provides a wrapper around a data array (\texttt{numpy} 
\texttt{ndarray}) and gives easy access to standard meta data in the header of 
the image, via the \texttt{.meta} attribute as well as some convienience 
properties for more standard meta data, \textit{i.e.} \texttt{.instrument}, 
\texttt{.wavelength} or \texttt{.coordinate\_system}.
The \texttt{Map} data type provides convenience methods for many functions 
such as, rotation and re-sampling as well as convenience visualisation 
functions, providing a easy to use as well as powerful interface.

The design of the map submodule is such that each instrument or 
detector can subclass the parent \texttt{GenericMap} class to implement 
special meta data handling or other data specific functions. Each subclass 
of \texttt{GenericMap} can register with the \texttt{Map} factory class and 
by implementing a method that returns \texttt{True} if the meta data 
matches meta data for that instrument or detector, the \texttt{Map} factory 
will automatically return an instance of the specific \texttt{GenericMap} 
subclass. As of version 0.4 SunPy has \texttt{Map} specialisations for the 
following instruments: \textit{IRIS} SJI frames, \textit{SDO/AIA} and 
\textit{HMI}, \textit{SOHO/EIT} and	\textit{LASCO}, \textit{STEREO/EUVI} and 
\textit{COR}, \textit{YOHKOH/SXT}, \textit{RHESSI}, \textit{PROBA2/SWAP} and 
\textit{HINODE/XRT}.

As can been seen in Listing \ref{code:aia_1} the map datatype parses header 
information for many purposes including visualisation on correct physical axes 
and limb plotting as well as making this information easily available to the 
user.

\begin{listing}[H]
\begin{minted}{pycon}

@@

>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap
SunPy AIAMap
---------
Observatory: SDO
Instrument:  AIA
Detector:    AIA
Measurement: 171
Obs Date:    2011-03-19T10:54:00.34
dt:          1.999601
Dimension:   [1024, 1024]
[dx, dy] =   [2.400000, 2.400000]

@@

>>> aia.peek()
\end{minted}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\caption{Demonstration of the \texttt{AIAMap} specilisation of 
\texttt{GenericMap}. The Map is created from a \textit{AIA} FITS file and the 
key meta data and array overview is printed. Then a quick view plot is created 
by using the \texttt{.peek()} method.}
\label{code:aia_1}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}{python}
import matplotlib.pyplot as plt
import sunpy.map

@@

\begin{listing}[H]
\begin{minted}{python}
import sunpy.map

### 2014-01-21 2-2-Lightcurve.tex
add stuff to dataypes

A \textit{Lightcurve} object may be created using a range of different methods. Firstly, a \textit{Lightcurve} may be created for a specific instrument based on an input time range (see example below). In this example, the Lightcurve constructor searches remote sites for the GOES X-ray data specified by the time interval, downloads the required files and subsequently creates the object.

\begin{lstlisting}[language=Python]
from sunpy import lightcurve
goes=lightcurve.GOESLightCurve.create(`2011-06-07 06:00',
`2011-06-07 08:00')
goes.peek()

\end{lstlisting}

Alternatively, if the data file already exists on the local system, the \textit{Lightcurve} object may instead be initialized using that file as input (see example below).

\begin{lstlisting}[language=Python]
from sunpy import lightcurve
file=`lyra_20110607-000000_lev2_std.fits'
lyra=lightcurve.LYRALightCurve.create(file)
lyra.peek()

\end{lstlisting}

>>>>

\subsubsection{Creating and using Lightcurves}

A \textit{Lightcurve} object may be created using a range of different methods. 
Firstly, a \textit{Lightcurve} may be created for a specific instrument based 
on an input time range (see Listing \ref{goes_lc_code} below). In this example, 
the Lightcurve constructor searches remote sites for the GOES X-ray data 
specified by the time interval, downloads the required files and subsequently 
creates the object (see also Listing \ref{goes_lc_code}).

\begin{listing}[H]
\begin{minted}{python}
>>> from sunpy import lightcurve
>>> goes=lightcurve.GOESLightCurve.create('2011-06-07 06:00',
'2011-06-07 08:00')
>>> goes.peek()
\end{minted}
\includegraphics[width=14cm]{goes_lightcurve.pdf}
\label{goes_lc_code}
\caption{Creating a GOES lightcurve for the time interval 06:00 - 08:00 UT on 
2011 June 7 using a time range, and the result of the \texttt{peek()} command.}
\end{listing}

Alternatively, if the data file already exists on the local system, the 
\textit{Lightcurve} object may instead be initialized using that file as input 
(see Listing \ref{lyra_lc_code}).

\begin{listing}[H]
\begin{minted}{python}
>>> from sunpy import lightcurve
>>> file='lyra_20110607-000000_lev2_std.fits'
>>> lyra=lightcurve.LYRALightCurve.create(file)
>>> lyra.peek()
\end{minted}
\includegraphics[width=14cm]{lyra_lightcurve.pdf}
\label{lyra_lc_code}
\caption{Creating a LYRA lightcurve for the full day of 2011 June 7 using an 
existing data file, and the result of the \texttt{peek()} command}
\end{listing}

Once the \textit{Lightcurve} has been created it may be manipulated in a variety of ways in order to perform time series analysis. Continuing from the above, we show a few simple examples of using a LYRA lightcurve.

In the above case (Listing \ref{lyra_lc_code}), we created a LYRA Lightcurve spanning the full day of 2011 June 7. To extract the sub-interval 06:00 - 07:00 UT into a new \textit{Lightcurve} called \verb|lyra_zoom| we simply do the following:

\begin{listing}[H]
\begin{minted}{python}
>>> from sunpy.time import TimeRange
>>> tran=TimeRange('2011-06-07 06:00','2011-06-07 07:00')
>>> lyra_zoom = lyra.truncate(tran)
\end{minted}
\label{lyra_truncate}
\caption{Extracting a sub-interval from a Lightcurve.}
\end{listing}

Also in the example shown in Listing \ref{lyra_lc_code}, direct access to the data and time information in the object is available through \verb|lyra.data| and \verb|lyra.data.index| respectively.

\begin{listing}[H]
\begin{minted}{python}
>>> print lyra.data
<class 'pandas.core.frame.DataFrame'>
DatetimeIndex: 1496629 entries, 2011-01-28 00:00:00.110000 to 
2011-01-28 23:59:59.983993
Data columns (total 4 columns):
CHANNEL1    1496629  non-null values
CHANNEL2    1496629  non-null values
CHANNEL3    1496629  non-null values
CHANNEL4    1496629  non-null values
dtypes: float64(4)

>>> print lyra.data.index
<class 'pandas.tseries.index.DatetimeIndex'>
[2011-01-28 00:00:00.110000, ..., 2011-01-28 23:59:59.983993]
Length: 1496629, Freq: None, Timezone: None
\end{minted}
\label{lyra_pandas}
\caption{Accessing the data and time axis in a Lightcurve}
\end{listing}

### 2014-01-21 2-Data_Types.tex
add stuff to dataypes

\section{Core Data Types}
One of the primary focus' of the SunPy library is to provide data structures 
that are 
specifically designed for the various types of data generally processed for 
solar physics. Currently \texttt{SunPy} provides three datatypes, each separate 
from one another to process 2D spatial data (Map), 1D temporal series 
(LightCurve) and 1 and 2D spectra (Spectrum and Spectra).

This section will give a brief overview of the \textit{current} functionality 
of each of these modules and then describe the development plan for future 
improvements to these submodules.

@@

This design philosophy makes the behaviour of SunPy's visualisation routines 
intuitive to those who already understand how matplotlib's interface is 
designed as well as allowing the use of the standard matplotlib commands to 
change the plot, \textit{e.g.} set the title.

>>>>

\section{Core Data Types}
One of the primary focus' of the SunPy library is to provide data structures 
that are specifically designed for the various types of data generally 
processed for solar physics. Currently SunPy provides three data 
types, each separate from one another to process 2D spatial data 
(\texttt{Map}), 1D temporal series (\texttt{LightCurve}) and 1 and 2D spectra 
(\texttt{Spectrum} and \texttt{Spectra}).

The design of SunPy's core data types is as a wrapper around a data object, 
either a \texttt{numpy} \texttt{ndarray} or a \texttt{pandas} 
\texttt{DataFrame}, for \texttt{Map} and \texttt{Spectra} and 
\texttt{LightCurve} respectively. Each data type has a \texttt{.data} attribute 
where this data object is located and some meta data. It is possible to 
instantiate the data types from various different sources, \textit{e.g.} files, 
urls, arrays or time ranges. The specific sources supported is dependant upon 
which data type and which instrument. All the SunPy data types support 
specialisation on a per-instrument basis, \textit{i.e.} \texttt{Map} has a 
\texttt{AIAMap} sub-type. This design allows simple extensions to cope with the 
diversity of data from different sources in solar physics. As well as support 
for different instruments and data sources each data type supports different 
visualisation wrappers that allow both quick view plotting (the peek command) 
and more fine-grained control plotting that integrates well with 
\texttt{matplotlib} (the plot command).

This section will give a brief overview of the \textit{current} functionality 
of each of these modules.

@@

This design philosophy makes the behaviour of SunPy's visualisation routines 
intuitive to those who already understand how matplotlib's interface is 
designed as well as allowing the use of the standard matplotlib commands to 
change the plot, \textit{e.g.} set the title. 
More technically what is done in the plotting routines is firstly, to make sure 
that the current state of the figure and axes is mantined in the same way as it 
is inside the \texttt{pyplot} functions. For \texttt{Map} this means setting 
\texttt{pyplot.sci()} or set current image, so that \texttt{plt.colorbar()} 
automatically detects the plotted \texttt{AxesImage} object. Secondly, it also 
means returning the same type of object as the \texttt{pyplot} interfaces 
function, \textit{i.e.} for \texttt{Map} it returns the \texttt{AxesImage} 
instance which is the same type of return as \texttt{pyplot.imshow}.
